### 2016-06-11 gruntz.tex
Fix mistakes in the Gruntz algorithm example
1. how to choose subexpression omega (this part was: wrong)
2. the rewrite stage (was: non-exsistent), I hope
now it's clear how to rewrite elements of the mrv set,
and this is not a magic step for our readers (or worse,
for "authors" of the SymPy) anymore.
3. mention the rewrite stage in comment about the
algorithm termination.
4. some minor fixes, i.e. "The goal is to calculate" was
dropped, the Gruntz algorithm IS about calculation limits
at the infinity.

SymPy calculates limits using the Gruntz algorithm, as described in%
~\cite{Gruntz1996limits}. The basic idea is as follows: any limit can be
converted to a limit $\lim\limits_{x\to\infty} f(x)$ by substitutions like
$x\to{1\over x}$. Then the subexpression $\omega$ (that converges
to zero as $x\to\infty$ faster than all other subexpressions) is identified in
$f(x)$, and $f(x)$ is expanded into a series with respect to $\omega$. Any
positive powers of $\omega$ converge to zero (while negative powers indicate
an infinite limit) and any constant term independent of
$\omega$ determines the limit. When a constant term still dependends on
$x$ the Gruntz algorithm is applied again until a final numerical value
is obtained as the limit.

To determine the most rapidly varying subexpression, the comparability classes
must first be defined, by calculating $L$:
\begin{equation}
L\equiv \lim_{x\to\infty} {\log |f(x)| \over \log |g(x)|}
\end{equation}
The relations $<$, $>$, and $\sim$ are defined as follows: $f>g$ when
$L=\pm\infty$ (it is said that $f$ is more rapidly varying than $g$, i.e., $f$
goes to $\infty$ or $0$ faster than $g$), $f<g$ when $L=0$ ($f$ is less
rapidly varying than $g$) and $f\sim g$ when $L\neq 0,\pm\infty$ (both $f$ and
$g$ are bounded from above and below by suitable integral powers of the
other). Note that if $f > g$, then $f > g^n$ for any $n$. Here
are some examples of comparability classes:
\[2 < x < e^x < e^{x^2} < e^{e^x}\]
\[2\sim 3\sim -5\]
\[x\sim x^2\sim x^3\sim {1\over x}\sim x^m\sim -x\]
\[e^x\sim e^{-x}\sim e^{2x}\sim e^{x+e^{-x}}\]
\[f(x)\sim{1\over f(x)}\]

The Gruntz algorithm is now illustrated with the following example:
\begin{equation}
    \label{gruntz_example_fn}
f(x) = e^{x+2e^{-x}} - e^x + {1\over x} \,.
\end{equation}
First, the set of most rapidly varying subexpressions is determined ---
the so-called \textit{mrv set}. For~\eqref{gruntz_example_fn}, the mrv set
$\{e^x, e^{-x}, e^{x+2e^{-x}}\}$ is obtained. These are all subexpressions of%
~\eqref{gruntz_example_fn} and they all belong to the same comparability
class. This calculation can be done using SymPy as follows:

% dict_keys output order varies
% no-doctest
\begin{verbatim}
>>> from sympy.series.gruntz import mrv
>>> mrv(exp(x+2*exp(-x))-exp(x) + 1/x, x)[0].keys()
dict_keys([exp(x + 2*exp(-x)), exp(x), exp(-x)])
\end{verbatim}

Next, an arbitrary item $\omega$ is taken from mrv set that converges to zero for
$x\to\infty$ and doesn't have subexpressions in the given mrv set.  In our case,
only item $\omega=e^{-x}$ can be accepted.  If such a term is not
present in the mrv set (i.e., all terms converge to infinity instead of zero),
the relation $f(x)\sim {1\over f(x)}$ can be used.

The next step is to rewrite the mrv set in terms of $\omega=g(x)$.  Every element
$f(x)$ of the mrv set is rewritten as $A \omega^c$, where
\begin{equation}
\label{gruntz_rewrite}
c = \lim\limits_{x\to\infty} \frac{\log{f(x)}}{\log{g(x)}},
\qquad
A = e^{\log f - c \log g}
\end{equation}
Note that this step include calculation of more simple limits, for instance
\begin{equation}
	\lim\limits_{x\to\infty} \frac{\log{e^{x + 2 e^{-x}}}}{\log e^{-x}}=
	\lim\limits_{x\to\infty} \frac{x + 2 e^{-x}}{-x} = -1
\end{equation}
In our example we obtain the rewritten mrv set: $\{{1\over\omega},
\omega, {1\over\omega}e^{2\omega}\}$. This can be done in the SymPy with
\begin{verbatim}
>>> from sympy.series.gruntz import mrv, rewrite
>>> m = mrv(exp(x+2*exp(-x))-exp(x) + 1/x, x)
>>> w = Symbol('w')
>>> rewrite(m[1], m[0], x, w)[0]
     2*w
1   e      1
- + ---- - -
x    w     w
\end{verbatim}
Then the rewritten subexpressions are
substituted back into $f(x)$ in~(\ref{gruntz_example_fn})
and the result is expanded with respect to~$\omega$:
\begin{equation}
    \label{gruntz_example_fn2}
f(x) = {1\over x}-{1\over\omega}+{1\over\omega}e^{2\omega}
     = 2+{1\over x} + 2\omega + O(\omega^2)
\end{equation}

Since $\omega$ is from the mrv set, then in the limit as $x\to\infty$,
$\omega\to0$, and so $2\omega + O(\omega^2) \to 0$ in~\eqref{gruntz_example_fn2}:
\begin{equation}
f(x) = {1\over x}-{1\over\omega}+{1\over\omega}e^{2\omega}
    = 2+{1\over x} + 2\omega + O(\omega^2)
    \to 2 + {1\over x}
\end{equation}

In our example the result ($2+{1\over x}$) still depends on $x$, so the above procedure is
repeated until just a value independent of $x$ is obtained. This is the final
limit. In the above case the limit is $2$, as can be
verified by SymPy:

\begin{verbatim}
>>> limit(exp(x+2*exp(-x))-exp(x) + 1/x, x, oo)
2
\end{verbatim}

In general, when $f(x)$ is expanded in terms of $\omega$, the following is obtained:
\begin{equation}
f(x) = \underbrace{O\left({1\over \omega^3}\right)}_\infty
    + \underbrace{C_{-2}(x)\over \omega^2}_\infty
    + \underbrace{C_{-1}(x)\over \omega}_\infty
    + {C_{0}(x)}
    + \underbrace{C_{1}(x)\omega}_0
    + \underbrace{O(\omega^2)}_0
\end{equation}
The positive powers of $\omega$ are zero. If there are any negative powers of
$\omega$, then the result of the limit is infinity, otherwise the limit is
equal to $\lim\limits_{x\to\infty} C_0(x)$. The expression $C_0(x)$ is always
simpler than original $f(x)$, same is true for limits, arising in the
rewrite stage~(\ref{gruntz_rewrite}), so the algorithm converges. A proof of this and further
details on the algorithm are given in Gruntz's PhD thesis~\cite{Gruntz1996limits}.