### 2016-06-11 supplement.tex
Remove custom modifications from wlpeerj.cls

%% Submissions for peer-review must enable line-numbering 
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need 
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file.

\documentclass[fleqn,10pt,lineno,numbers]{wlpeerj} % for journal submissions
% \documentclass[fleqn,10pt]{wlpeerj} % for preprint submissions

%\usepackage[numbers]{natbib}

\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[scaled=0.8]{DejaVuSansMono}

\usepackage{hyperref}
\usepackage{graphicx}
\usepackage[all]{xy}
\usepackage{amsmath}
\usepackage{caption}
\graphicspath{ {images/} }

% Makes quote characters in monospace font not be curly
\usepackage{upquote}

\usepackage{amsmath}
\usepackage{url}
\usepackage{hyperref}

% this is required for all the \url{} commands in the bib file
%\usepackage{hyperref}

% for nice units
\usepackage{siunitx}

% for images: png, pdf, etc
\usepackage{graphicx}

% for nice table formatting, i.e., /toprule, /midrule, etc
\usepackage{booktabs}

% to allow for \verb++ declarations in captions.
\usepackage{cprotect}

% to allow usage of \mathbb symbols
\usepackage{amssymb}

\usepackage{longtable}

\usepackage{listings}

\begin{document}

\flushbottom
\thispagestyle{empty}%
\vskip-36pt%
{\raggedright\sffamily\bfseries\fontsize{20}{25}\selectfont SymPy: Symbolic Computing in Python\par}%
\vskip10pt
{\raggedright\sffamily\fontsize{12}{16}\selectfont  Supplementary material\par}
\vskip25pt%

As in the paper, all examples in the supplement assume that the following
has been run:
\begin{verbatim}
>>> from sympy import *
>>> x, y, z = symbols('x y z')
\end{verbatim}

\section{Limits: The Gruntz Algorithm}

\input{gruntz.tex}

% Series module (Formal Power Series, Fourier Series)
\section{Series}

\input{series.tex}

% Logic module
\section{Logic}

\input{logic}

\section{Diophantine Equations}

\input{diophantine}

% Sets
\section{Sets}
\input{sets}

\section{Category Theory}
\input{categories}

\section{SymPy Gamma}\label{sympy-gamma}

\input{gamma}

\section{SymPy Live}\label{sympy-live}

\input{live}

\section{Comparison with Mathematica}

\input{comparison_with_mma}

\section{Other Projects that use SymPy}

\input{other_projects_that_use_sympy}

\section{Tensors}

\input{tensors}

\section{Numerical simplification}

\input{nsimplify}

\section{Examples}

\input{examples}

\section{References}

\bibliography{paper}

\end{document}