### 2016-08-16 architecture.tex
Address questions 16 and 17 of reviewer 3
footnote was moved above, to the section 2.2.

Software architecture is of central importance in any large software
project because it establishes predictable patterns of usage and
development~\cite{Shaw1996}. This section describes the essential
structural components of SymPy, provides justifications for the design
decisions that have been made, and gives example
user-facing code as appropriate.

\subsection{Basic Usage}

% symbols, various ways to declare them

Symbolic variables, called symbols, must be defined and assigned to
Python variables before they can be used. This is typically done through the
\texttt{symbols} function, which may create multiple symbols in a single
function call. For instance,
\begin{verbatim}
>>> x, y, z = symbols('x y z')
\end{verbatim}
creates three symbols representing variables named $x$, $y$, and $z$. In this
particular instance, these symbols are all assigned to Python variables of the
same name. However, the user is free to assign them to different
Python variables, while representing the same symbol, such as
\texttt{a, b, c = symbols(\textquotesingle{}x y z\textquotesingle{})}.
In order to minimize potential confusion, though, all examples in this paper will
assume that
the symbols \verb|x|, \verb|y|, and \verb|z| have been assigned to Python variables
identical to their symbolic names.

Expressions are created from symbols using Python's mathematical syntax.  For
instance, the following Python code creates the expression $(x^2 - 2x + 3)/y$.
Note that the expression remains unevaluated: it is represented symbolically.

\begin{verbatim}
>>> (x**2 - 2*x + 3)/y
(x**2 - 2*x + 3)/y
\end{verbatim}

Importantly, SymPy expressions are immutable. This simplifies the design of
SymPy by allowing expression interning. It also enables expressions to be
hashed, that is used to implement caching in SymPy.

%% I volunteer to write this section. --Aaron
%%
%% Representing symbolic expressions using Python objects

\subsection{The Core}

A computer algebra system stores mathematical expressions as data
structures.  For example, the mathematical expression $x + y$ is represented as
a tree with three nodes, $+$, $x$, and $y$, where $x$ and $y$ are ordered
children of $+$.  As users manipulate
mathematical expressions with traditional mathematical syntax, the CAS
manipulates the underlying data structures.  Automated optimizations and
computations such as integration, simplification, etc.\ are all functions that
consume and produce expression trees.

In SymPy every symbolic expression is an instance of a Python \texttt{Basic}
class\footnote{Some
  internal classes, such as those used in the polynomial module, do not follow
  this rule for efficiency reasons.}, a superclass of all SymPy types providing
common methods to all SymPy tree-elements, such as traversals.  The children of a node in the
tree are held in the \texttt{args} attribute.  A terminal or leaf node in the
expression tree has empty \texttt{args}.

For example, consider the expression $xy + 2$:
\begin{verbatim}
>>> expr = x*y + 2
\end{verbatim}
By order of operations, the parent of the expression tree for \texttt{expr} is
an addition, so it is of type \texttt{Add}. The child nodes of \texttt{expr} are
\texttt{2} and \texttt{x*y}.
\begin{verbatim}
>>> type(expr)
<class 'sympy.core.add.Add'>
>>> expr.args
(2, x*y)
\end{verbatim}

Descending further down into the expression tree yields the full expression. For
example, the next child node (given by \texttt{expr.args[0]}) is
\texttt{2}. Its class is \texttt{Integer}, and it has an empty \texttt{args}
tuple, indicating that it is a leaf node.
\begin{verbatim}
>>> expr.args[0]
2
>>> type(expr.args[0])
<class 'sympy.core.numbers.Integer'>
>>> expr.args[0].args
()
\end{verbatim}

A useful way to view an expression tree is using the \texttt{srepr} function, which
returns a string representation of an expression as valid Python code\footnote{
The \texttt{dotprint} function from the \texttt{sympy.printing.dot} package
prints output to dot format, which can be rendered with Graphviz to
visualize expression trees graphically.}
with all the nested class constructor calls to create the given expression.
\begin{verbatim}
>>> srepr(expr)
"Add(Mul(Symbol('x'), Symbol('y')), Integer(2))"
\end{verbatim}

Every SymPy expression satisfies a key identity invariant:
\begin{verbatim}
expr.func(*expr.args) == expr
\end{verbatim}
This means that expressions are
rebuildable from their \texttt{args}.\footnote{\texttt{expr.func} is used
instead of \texttt{type(expr)} to allow the function of an expression to be
distinct from its actual Python class. In most cases the two are the same.}
Note that in SymPy the \texttt{==} operator represents exact
structural equality, not mathematical equality. This allows testing if any two
expressions are equal to one another as expression trees. For example, even
though ${(x + 1)}^2$ and $x^2 + 2x + 1$ are equal mathematically, SymPy gives
\begin{verbatim}
>>> (x + 1)**2 == x**2 + 2*x + 1
False
\end{verbatim}
because they are different as expression trees (the former is a \verb|Pow|
object and the latter is an \verb|Add| object).

Python allows classes to override mathematical operators. The Python
interpreter translates the above \texttt{x*y + 2} to, roughly,
\verb|(x.__mul__(y)).__add__(2)|. Both \texttt{x} and \texttt{y}, returned
from the \texttt{symbols} function, are \texttt{Symbol} instances. The
\texttt{2} in the expression is processed by Python as a literal, and is
stored as Python's built in \texttt{int} type. When \texttt{2} is passed to the
\verb|__add__| method of \texttt{Symbol}, it is converted to the SymPy type
\verb|Integer(2)| before being stored in the resulting expression tree. In
this way, SymPy expressions can be built in the natural way using Python
operators and numeric literals.

%%
%% Assumptions
\subsection{Assumptions}
\label{sec:assumptions}

SymPy performs logical inference through its assumptions system. The
assumptions system allows users to specify that symbols have certain common
mathematical properties, such as being positive, imaginary, or integral. SymPy
is careful to never perform simplifications on an expression unless the
assumptions allow them. For instance, the identity $\sqrt{t^2} = t$ holds if
$t$ is nonnegative ($t\ge 0$). However, for general complex $t$, no
such identity holds.

By default, SymPy performs all calculations assuming that symbols are
complex valued. This assumption makes it easier to treat mathematical problems
in full generality.
\begin{verbatim}
>>> t = Symbol('t')
>>> sqrt(t**2)
sqrt(t**2)
\end{verbatim}

By assuming the most general case, that \texttt{t} is complex by default, SymPy
avoids performing mathematically invalid operations. However, in many cases
users will wish to simplify expressions containing terms like $\sqrt{t^2}$.

Assumptions are set on \texttt{Symbol} objects when they are created. For
instance \texttt{Symbol(\textquotesingle{}t\textquotesingle{}, positive=True)} will create a symbol named
\texttt{t} that is assumed to be positive.
\begin{verbatim}
>>> t = Symbol('t', positive=True)
>>> sqrt(t**2)
t
\end{verbatim}

Some of the common assumptions that SymPy allows are \texttt{positive},
\texttt{negative}, \texttt{real}, \texttt{nonpositive}, \texttt{integer},
\texttt{prime} and \texttt{commutative}\footnote{SymPy assumes that two
  expressions $A$ and $B$ commute with each other multiplicatively, that is,
  $A\cdot B=B\cdot A$, unless they both have \texttt{commutative=False}. Many
  algorithms in SymPy require special consideration to work correctly with
  noncommutative products.}. Assumptions on any object can be checked with the
\verb|is_|\texttt{\textit{assumption}} attributes, like \verb|t.is_positive|.

Assumptions are only needed to restrict a domain so that certain
simplifications can be performed. They are not required to make the domain match
the input of a function. For instance, one can create the object
$\sum_{n=0}^m f(n)$ as \verb|Sum(f(n), (n, 0, m))| without setting
\texttt{integer=True} when creating the Symbol object \texttt{n}.

The assumptions system additionally has deductive capabilities. The
assumptions use a three-valued logic using the Python built in objects
\texttt{True}, \texttt{False}, and \texttt{None}. Note that \texttt{False}
is returned if the SymPy object doesn't or can't have the assumption.
For example, both \texttt{I.is\_real} and \texttt{I.is\_prime} return
\texttt{False} for the imaginary unit \texttt{I}.

\texttt{None} represents the
``unknown'' case. This could mean that given assumptions do not unambiguously
specify the truth of an attribute. For instance,
\texttt{Symbol(\textquotesingle{}x\textquotesingle{}, real=True).is\_positive}
will give \texttt{None} because a real symbol might be positive or negative.
The \texttt{None} could also mean that not enough is known or implemented to
compute the given fact. For instance, \texttt{(pi + E).is\_irrational} gives
\texttt{None}, because determining whether $\pi + e$ is rational or irrational
is an open problem in mathematics~\cite{lang1966introduction}.

Basic implications between the facts are used to deduce assumptions. For
instance, the assumptions system knows that being an integer implies being
rational.
\begin{verbatim}
>>> i = Symbol('i', integer=True)
>>> i.is_rational
True
\end{verbatim}
Furthermore, expressions compute the assumptions on themselves
based on the assumptions of their arguments. For instance, if \texttt{x} and
\texttt{y} are both created with \texttt{positive=True}, then
\texttt{(x + y).is\_positive} will be \texttt{True} whereas
\texttt{(x - y).is\_positive} will be \texttt{None}.

%% TODO: describe how assumptions are implemented

%%
%% Extensibility
\subsection{Extensibility}

While the core of SymPy is relatively small, it has been extended to a wide variety
of domains by a broad range of contributors.
This is due, in part, to the fact that the same language, Python,
is used both for the internal implementation and the external usage by users.
All of the extensibility capabilities available to
users are also utilized by SymPy itself. This eases the transition pathway from
SymPy user to SymPy developer.

The typical way to create a custom SymPy object is to subclass an existing
SymPy class, usually  \texttt{Basic}, \texttt{Expr}, or
\texttt{Function}. As it was stated before, almost all SymPy classes used for
expression trees should be subclasses of the base class \texttt{Basic}.
\texttt{Expr} is the subclass for mathematical expressions (including \texttt{Symbol}) that can be
added and multiplied together. Instances of \texttt{Expr} typically represent
complex numbers, but may also include other ``rings'' like matrix expressions.
Not all SymPy classes are subclasses of \texttt{Expr}. For instance, logic expressions such
as \verb|And(x, y)| are subclasses of \texttt{Basic} but not of \texttt{Expr}.

The \texttt{Function} class is a subclass of \texttt{Expr} which makes it
easier to define mathematical functions called with arguments. This includes
named functions like $\sin(x)$ and $\log(x)$ as well as undefined functions
like $f(x)$. Subclasses of \texttt{Function} should define a
class method \texttt{eval}, which returns a canonical form of the function
application (usually an instance of some other class, i.e. a \texttt{Number})
or \texttt{None}, if for given arguments that function should not be
automatically evaluated.

Many SymPy functions perform various evaluations down the expression tree.
Classes define their behavior in such functions by defining a relevant
\verb|_eval_|\texttt{\textit{*}} method. For instance, an object can indicate
to the \texttt{diff} function how to take the derivative of itself by defining
the \verb|_eval_derivative(self, x)| method, which may in turn call
\texttt{diff} on its \texttt{args}. (Subclasses of \texttt{Function} should
implement \texttt{fdiff} method instead, it returns the derivative of the function
without considering the chain rule.) The most common
\verb|_eval_|\texttt{\textit{*}} methods relate to the assumptions:
\verb|_eval_is_|\texttt{\textit{assumption}} is used to deduce
\textit{assumption} on the object.

As an example of the notions presented in this section,
Listing~\ref{fig:gamma-example} presents a minimal version of the gamma
function $\Gamma(x)$ from SymPy, which evaluates itself on positive integer
arguments, has the positive and real assumptions defined, can be rewritten in
terms of factorial with \verb|gamma(x).rewrite(factorial)|, and can be
differentiated. \texttt{self.func} is used throughout
instead of referencing \texttt{gamma} explicitly so that potential subclasses
of \texttt{gamma} can reuse the methods.

\lstset{
  basicstyle=\ttfamily,
}

\begin{lstlisting}[caption={A minimal implementation of \texttt{sympy.gamma}.},label=fig:gamma-example]
from sympy import Integer, Function, floor, factorial, polygamma

class gamma(Function)
    @classmethod
    def eval(cls, arg):
        if isinstance(arg, Integer) and arg.is_positive:
            return factorial(arg - 1)

    def _eval_is_positive(self):
        x = self.args[0]
        if x.is_positive:
            return True
        elif x.is_noninteger:
            return floor(x).is_even

    def _eval_is_real(self):
        x = self.args[0]
        # noninteger means real and not integer
        if x.is_positive or x.is_noninteger:
            return True

    def _eval_rewrite_as_factorial(self, z):
        return factorial(z - 1)

    def fdiff(self, argindex=1):
        from sympy.core.function import ArgumentIndexError
        if argindex == 1:
            return self.func(self.args[0])*polygamma(0, self.args[0])
        else:
            raise ArgumentIndexError(self, argindex)
\end{lstlisting}
The gamma function implemented in SymPy has many more capabilities than the
above listing, such as evaluation at rational points and series expansion.