### 2016-08-22 acknowledgements.tex
Reorganize the acknowledgements.tex
We only thank the institutions, but do not explain what they are, per
reviewer's comment.

% People should add their funding agency acknowledgements here.
All authors thank the Google Summer of Code for its financial support of
students who contributed to SymPy.

The author of this paper Ondřej Čertík thanks the Los Alamos National
Laboratory for its financial support.

The author of this paper Richard P. Muller thanks Sandia National Laboratory
for its financial support.

The author of this paper Francesco Bonazzi thanks the Deutsche
Forschungsgemeinschaft (DFG) for its financial support via the International
Research Training Group 1524 "Self-Assembled Soft Matter Nano-Structures at
Interfaces."