### 2016-08-22 comparison_with_mma.tex
CASs -> CASes
Reviewer 3, comment 3.

Wolfram Mathematica is a popular proprietary CAS that
features highly advanced algorithms, has a core written in
C++~\cite{Wolfram2003book}, and interprets its own programming language,
Wolfram Language.

% M-expressions

Analogous to Lisp S-expressions,
Mathematica uses its own style of M-expres\-sions,
which are arrays of either atoms or other M-expressions.
The first element of the expression identifies the type of the expression
and is indexed by zero, and the first argument is indexed starting with one.
In SymPy, expression arguments are stored in a Python tuple
(that is, an immutable array),
while the expression type is identified by the type of the object storing the
expression.

% Attributes

Mathematica can associate attributes to its atoms.
Attributes may define mathematical properties and behavior of the nodes
associated to the atom.
In SymPy, the usage of static class fields is roughly similar to Mathematica's
attributes, though other programming patterns may also be used to achieve an
equivalent behavior such as class inheritance.

% Expression mutability

Unlike SymPy, Mathematica's expressions are mutable:
one can change parts of the expression tree without the need of
creating a new object.
The mutability of Mathematica expressions allows for a lazy updating of any references
to a given data structure.

% * comparison with Mathematica: commutativity, associative expressions, one-identity. Advantage of SymPy: multiplicative commutativity defined on symbols.
% Products and commutativity

Products in Mathematica are determined by some built in node types, such as
\texttt{Times}, \texttt{Dot}, and others.  \texttt{Times} is a representation of
the \texttt{*} operator, and is always meant to represent a commutative product
operator.  The other notable product is \texttt{Dot}, which represents the
\texttt{.} operator.  This product represents matrix multiplication. It is not
commutative.  Unlike
Mathematica, SymPy determines commutativity with respect to multiplication from
the expression type of the factors.  Mathematica puts the \texttt{Orderless} attribute
on the expression type.

% Associative expressions.

Regarding associative expressions,
SymPy handles associativity of sums and products by automatically flattening them,
Mathematica specifies the \texttt{Flat} attribute on the expression type.

% One identity

% Pattern matching

Mathematica relies heavily on pattern matching---even the so-called equivalent of function declaration is in reality
the definition of a pattern generating an expression tree transformation
on input expressions.
%
Mathematica's pattern matching is sensitive to
associative, commutative, and
one-identity properties of its expression tree
nodes.
%
SymPy has various ways to perform pattern matching.
All of them play a lesser role in the CAS than in Mathematica
and are basically available as a tool to rewrite expressions.
The differential equation solver in SymPy somewhat relies on pattern matching to
identify differential equation types, but it is envisaged to replace
that strategy with analysis of Lie symmetries in the future.
Mathematica's real advantage is the ability to add (at runtime) new overloading
to the expression builder or specific subnodes.
Consider for example:
\begin{verbatim}
In[1]:= Unprotect[Plus]
Out[1]= {Plus}

In[2]:= Sin[x_]^2 + Cos[y_]^2 := 1

In[3]:= x + Sin[t]^2 + y + Cos[t]^2
Out[3]= 1 + x + y
\end{verbatim}
This expression in Mathematica defines a substitution rule that overloads
the functionality of the \texttt{Plus} node (the node for additions in Mathematica).
A symbol with a trailing underscore is treated as a wildcard.
Although one may wish to keep this identity unevaluated, this example clearly
illustrates the potential to define
one's own immediate transformation rules.
In SymPy, the operations constructing the addition node in the expression tree
are Python class constructors
and cannot be modified at runtime.\footnote{Nonetheless, Python supports monkey
patching but it is a discouraged programming pattern.}
The way SymPy deals with extending the missing runtime overloadability functionality
is by subclassing the node types: subclasses may redefine the class constructor
to yield the proper extended functionality.

%% TODO list:
% * comparison with Mathematica: MatrixExp, product not always commutative, type inheritance (polymorphism) and advantage in unifying the product symbol * for symbols and matrices, pattern matching vs. single dispatch.

% Type inheritance and polymorphism

Unlike SymPy, Mathematica does not support type inheritance or poly\-morph\-ism~\cite{Fateman1992}.
% cite examples of class inheritance in SymPy:
%
SymPy relies heavily on class inheritance, but for the most part,
class inheritance is used to make sure that SymPy objects inherit the proper
methods and implement the basic hashing system.
%There are also cases where inheritance is used to extend the mathematical meaning of an expression.

% Matrices

While Mathematica interprets nested lists as matrices whenever
the sublists have the same length, matrices in SymPy are a type in their own
right, allowing ordinary operators and functions (like multiplication and
exponentiation) to be used as they traditionally are in mathematics.
\begin{verbatim}
>>> exp(Matrix([[1, 1],[0, 2]])) * Matrix([a, b])
Matrix([
[E*a + b*(-E + exp(2))],
[             b*exp(2)]])
\end{verbatim}

Using the standard multiplication in Mathematica performs an element-wise
product and
calling the exponential function \texttt{Exp} on a matrix
returns an element-wise exponentiation of its elements.

% * comparison with Mathematica: avoid misspelling variables through forced declaration (check that you can't do it in Mathematica).
% * evaluate=False vs HoldForm

Unevaluated expressions in Mathematica can be achieved in various ways,
most commonly with the \texttt{HoldForm} or \texttt{Hold} nodes,
that block the evaluation of subnodes by the parser.
Such a node cannot be expressed in Python because of greedy evaluation.
Whenever needed in SymPy, it is necessary to add the parameter \texttt{evaluate=False}
to all subnodes.

% * comparison with Mathematica: == is structural (not mathematical) equality

In Mathematica, the operator \texttt{==} returns a boolean whenever it is able
to immediately evaluate the truth of the equality, otherwise it returns an
\texttt{Equal} expression.  In SymPy, \texttt{==} means structural equality and
is always guaranteed to return a boolean expression.  To express a mathematical equality in
SymPy it is necessary to explicitly construct an instance of the \texttt{Equality}
class.

% * comparison with Mathematica: polynomial module.
% * comparison with Mathematica: space is product, ** vs ^

SymPy, in accordance with Python (and unlike the usual programming convention),
uses \texttt{**} to express the power operator, while Mathematica uses the more
common \verb|^|.

% * comparison with Mathematica: ( ) is Sequence, functions are generally uppercase.
% * comparsion with Mathematica: table of comparison?
% * comparison with Mathematica: Wolfram language has loads of operator overloading, functional paradigm.

SymPy's use of floating-point numbers is similar to that of most
other CASes, including Maple and Maxima.
By contrast, Mathematica uses a form
of significance arithmetic~\cite{Sofroniou2005precise} for approximate numbers.
This offers further protection against numerical errors,
although it comes with its own set of problems
(for a critique of significance arithmetic, see Fateman~\cite{Fateman1992}).
Internally, SymPy's \texttt{evalf} method works similarly to Mathematica's
significance arithmetic, but the semantics are isolated from the rest of the system.