### 2016-08-26 acknowledgements.tex
Fix some chktex issues

% People should add their funding agency acknowledgements here.
All authors thank the Google Summer of Code for its financial support of
students who contributed to SymPy.

The author of this paper Ondřej Čertík thanks the Los Alamos National
Laboratory for its financial support.

The author of this paper Richard P. Muller thanks Sandia National Laboratories
for their financial support.

The author of this paper Francesco Bonazzi thanks the Deutsche
Forschungsgemeinschaft (DFG) for its financial support via the International
Research Training Group 1524 ``Self-Assembled Soft Matter Nano-Structures at
Interfaces.''