### 2016-09-11 other_projects_that_use_sympy.tex
Clarify "usage" of Sympy: that use -> that depend on
Ondrej (and perhaps Aaron) won't mention the Diofant project
among SymPy's users, so we must clarify that we mean by "using".

There are several projects that depend on SymPy as a library for implementing
a part of their functionality. Some of them are listed below:

\begin{itemize}
\item
  \href{http://cadabra.science/index.html}{\textbf{Cadabra}}: Cadabra is
  a CAS designed specifically for the
  resolution of problems encountered in field theory.
\item
  \href{http://octave.sourceforge.net/symbolic/}{\textbf{Octave Symbolic}}:
  The Octave-Forge Symbolic package adds symbolic calculation features
  to GNU Octave. These include common CAS tools such
  as algebraic operations, calculus, equation solving, Fourier and
  Laplace transforms, variable precision arithmetic, and other features.
\item
  \href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}:
  Provides a Julia interface to SymPy using PyCall.
\item
  \href{https://mathics.github.io/}{\textbf{Mathics}}: Mathics is a
  free, general-purpose online CAS featuring Mathematica compatible
  syntax and functions. It is backed by highly extensible Python code,
  relying on SymPy for most mathematical tasks.
\item
  \href{http://mathpix.com/}{\textbf{Mathpix}}: An iOS App, that detects handwritten math as input, and uses
  SymPy Gamma to evaluate the math input and generate the relevant
  steps to solve the problem.
\item
  \href{http://openrave.org/docs/0.8.2/openravepy/ikfast/}{\textbf{IKFast}}:
  IKFast is a robot kinematics compiler provided by
  \href{http://openrave.org/}{OpenRAVE}. It analytically solves robot inverse
  kinematics equations and generates optimized C++ files. It uses SymPy for
  its internal symbolic mathematics.
\item
  \href{http://www.sagemath.org/}{\textbf{Sage}}: A CAS, visioned to be
  a viable free open source alternative to Magma, Maple, Mathematica and
  MATLAB\@. Sage includes many open source mathematical libraries, including
  SymPy.
\item
  \href{https://cloud.sagemath.com}{\textbf{SageMathCloud}}:
  SageMathCloud is a web-based cloud computing and course management
  platform for computational mathematics.
\item
  \href{http://www.pydy.org/}{\textbf{PyDy}}: Multibody Dynamics with
  Python.
\item
  \href{https://github.com/brombo/galgebra}{\textbf{galgebra}}:
  Geometric algebra (previously \texttt{sympy.galgebra}).
\item
  \href{http://yt-project.org/}{\textbf{yt}}: Python package for
  analyzing and visualizing volumetric data (\texttt{yt.units} uses SymPy).
\item
  \href{http://sfepy.org/}{\textbf{SfePy}}: Simple finite elements in
  Python, see section~\ref{sfepy}.
\item
  \href{http://quameon.sourceforge.net/}{\textbf{Quameon}}: Quantum
  Monte Carlo in Python.
\item
  \href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}:
  Experimental Python package for teaching linear circuit analysis.
\item
  \href{http://digitalcommons.calpoly.edu/cgi/viewcontent.cgi?article=1072\&context=physsp/}{\textbf{Quantum
  Programming in Python}}: Quantum 1D Simple Harmonic Oscillator and
  Quantum Mapping Gate.
\item
  \href{http://mech.fsv.cvut.cz/~stransky/software/latexexpr/doc/}{\textbf{LaTeX
  Expression project}}: Easy \LaTeX{} typesetting of algebraic expressions
  in symbolic form with automatic substitution and result computation.
\item
  \href{https://www.researchgate.net/publication/260585491_Symbolic_Statistics_with_SymPy/}{\textbf{Symbolic
  statistical modeling}}: Adding statistical operations to complex
  physical models.
\end{itemize}

\subsection{SfePy}\label{sfepy}

\href{http://sfepy.org/}{\textbf{SfePy}} (Simple finite elements in Python),
cf.~\cite{cimrman2014sfepy}, is a Python package for solving partial
differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
method~\cite{Zienkiewicz2013FEM}. SymPy is used within this package mostly for
code generation and testing, namely:

\begin{itemize}
\item generation of the hierarchical FE basis module, involving generation and
  symbolic differentiation of 1D Legendre and Lobatto polynomials, constructing
  the FE basis polynomials~\cite{Solin2003hpfem} and generating the C code;
\item generation of symbolic conversion formulas for various groups of elastic
  constants~\cite{Fung1993contmech}:  provide any two of the Young's modulus,
  Poisson's ratio, bulk modulus, Lam\'{e}'s first parameter, shear modulus
  (Lam\'{e}'s second parameter) or longitudinal wave modulus and get the other
  ones;
\item simple physical unit conversions, generation of consistent unit sets;
\item testing FE solutions using method of manufactured (analytical)
  solutions:  the differential operator of a PDE is symbolically applied and a symbolic
  right-hand side is created, evaluated in quadrature points, and subsequently
  used to obtain a numerical solution that is then compared to the analytical
  one;
\item testing accuracy of 1D, 2D and 3D numerical quadrature formulas
  (cf.~\cite{Abramowitz1964handbook}) by generating polynomials of suitable
  orders, integrating them, and comparing the results with those obtained by
  the numerical quadrature.
\end{itemize}