### 2016-09-27 projects_that_depend_on_sympy.tex
Compactify projects table
Domain column was removed.  I believe it can't be very usefull.  E.g.
for most entries there should be something like "CAS".  Free text
in the Description should be enough.

There are several projects that depend on SymPy as a library for implementing
a part of their functionality. A selection of these projects are listed in
Table~\ref{projects-table}.

\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.8\linewidth}}
\caption{Selected projects that depend on SymPy.\label{projects-table}}\\
\toprule
\textbf{Project name} & \textbf{Description} \\
\midrule

\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
  analog of the Wolfram|Alpha.  See more in the supplementary material. \\

\href{http://cadabra.science/index.html}{\textbf{Cadabra}}~\cite{Peeters2007cadabra} &
  CAS designed specifically for the resolution of problems
  encountered in field theory. \\

\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}}~\cite{OctSymPy} &
  An implementation of a symbolic toolbox for Octave using SymPy. \\

\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}}~\cite{SymPy.jl} &
  Provides a Julia interface to SymPy using PyCall. \\

\href{https://mathics.github.io/}{\textbf{Mathics}}~\cite{Mathics} &
  A free, online CAS featuring Mathematica compatible
  syntax and functions. \\

\href{http://mathpix.com/}{\textbf{Mathpix}}~\cite{Mathpix} & An iOS App, that detects handwritten math as input, and uses
  SymPy Gamma to evaluate the math input and generate the relevant
  steps to solve the problem. \\

\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}}~\cite{diankov2010ikfast} &
  IKFast is a robot kinematics compiler provided by
  \href{http://openrave.org/}{OpenRAVE}. \\

\href{http://www.sagemath.org/}{\textbf{SageMath}}~\cite{sagemath} &
  A free open-source mathematics software system, builds on top of many
  existing open-source packages, including SymPy. \\

\href{http://www.pydy.org/}{\textbf{PyDy}}~\cite{gede2013constrained} & Multibody Dynamics with
  Python. \\

\href{https://github.com/brombo/galgebra}{\textbf{galgebra}}~\cite{galgebra} &
  A module for geometric algebra (previously \texttt{sympy.galgebra}). \\

\href{http://yt-project.org/}{\textbf{yt}}~\cite{2011ApJS..192....9T} & Python package for
  analyzing and visualizing volumetric data. \\

\href{http://sfepy.org/}{\textbf{SfePy}}~\cite{cimrman2014sfepy} &
  A Python package for solving partial
  differential equations (PDEs) in 1D, 2D and 3D by the finite element (FE)
  method~\cite{Zienkiewicz2013FEM}. \\

\href{http://quameon.sourceforge.net/}{\textbf{Quameon}}~\cite{quameon} & Quantum
  Monte Carlo in Python. \\

\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}}~\cite{lcapy} &
  Experimental Python package for teaching linear circuit analysis. \\
\bottomrule
\end{longtable}