### 2016-09-27 coverletter.tex
Split the cover letter into a separate document so we can us the letter class

\documentclass{letter}
\usepackage{hyperref}
\signature{Aaron Meurer}
\address{address}
\begin{document}

\begin{letter}{address}
\opening{Dear Editors,}

Dear Editors, \bigskip

We thank you and the referees for your thoughtful comments and suggestions on
our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
address your comments and concerns point by point below. Below we give your
and the reviewers' comments on white background and our replies on gray
background. \bigskip

Reviewer 1 had concerns that the paper was out of scope, but the editor has
confirmed that it is in scope. Many of the points that reviewer 3 raised were
in regards to the relevance of various sections. The intended scope of this
paper was to be an architecture paper for SymPy's 1.0 release. We have
examined reviewer 3's points through this lens. In light of this, which
contents were previously in the supplement vs.\ the main manuscript may have
been confusing to a general audience. To address this, we have restructured
the manuscript to make the paper more readable to those not already experts in
SymPy or symbolic mathematics. \bigskip

We thank you again for your consideration. Given these edits, we believe that
the manuscript is now suitable for publication in PeerJ Computer Science. If
there are further questions or concerns do not hesitate to contact us. \bigskip

Best regards,\bigskip

\end{letter}
\end{document}