### 2016-09-28 coverletter.tex
Add the sentence back in
It was removed in 59dcf547dfd813fe91a377c6ddcfda6462f66dbf.

\documentclass{letter}

\usepackage{graphicx}

\usepackage{hyperref}
% \signature{Aaron Meurer}
\address{Aaron Meurer \\
ERGS: \url{http://www.ergs.sc.edu/} \\
Nuclear Eng., Mechanical Eng. Dept. \\
University of South Carolina \\
541 Main Street, St. 011 \\
Columbia, SC 29201 \\
\url{asmeurer@gmail.com}}

\date{September 31, 2016} % stand-in
\begin{document}

\begin{letter}{}
\opening{Dear Editors,}

We thank you and the referees for your thoughtful comments and suggestions on
our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
address your comments and concerns point by point below. Below we give your
and the reviewers' comments on white background and our replies on gray
background.

Reviewer 1 had concerns that the paper was out of scope, but the editor has
confirmed that it is in scope. Many of the points that reviewer 3 raised were
in regards to the relevance of various sections. The intended scope of this
paper was to be an architecture paper for SymPy's 1.0 release. We have
examined reviewer 3's points through this lens. In light of this, which
contents were previously in the supplement vs.\ the main manuscript may have
been confusing to a general audience. To address this, we have restructured
the manuscript to make the paper more readable to those not already experts in
SymPy or symbolic mathematics.

We thank you again for your consideration. Given these edits, the manuscript
has been greatly improved and we believe that it is now suitable for
publication in PeerJ Computer Science. If there are further questions or
concerns please do not hesitate to contact us.

% Main text
\closing{Best regards,\\
\fromsig{\includegraphics[scale=0.07]{aaron-signature.png}} \\
\fromname{Aaron Meurer}
}

\end{letter}
\end{document}