### 2016-09-28 projects_that_depend_on_sympy.tex

There are several projects that depend on SymPy as a library for implementing
a part of their functionality. A selection of these projects are listed in
Table~\ref{projects-table}.

\begin{longtable}[htbc]{>{\raggedright}p{0.2\linewidth}p{0.74\linewidth}}
\caption{Selected projects that depend on SymPy.\label{projects-table}}\\
\toprule
\textbf{Project name} & \textbf{Description} \\
\midrule

\href{http://sympygamma.com/}{\textbf{SymPy Gamma}} & An open source
  analog of Wolfram|Alpha that uses SymPy~\cite{SymPyGamma}.  There is more
  information about SymPy Gamma supplementary material. \\

\href{http://cadabra.science/index.html}{\textbf{Cadabra}} &
  A CAS designed specifically for the resolution of problems
  encountered in field theory~\cite{Peeters2007cadabra}. \\

\href{https://github.com/cbm755/octsympy}{\textbf{GNU Octave Symbolic Package}} &
  An implementation of a symbolic toolbox for Octave using SymPy~\cite{OctSymPy}. \\

\href{https://github.com/jverzani/SymPy.jl}{\textbf{SymPy.jl}} &
  A Julia interface to SymPy, provided using PyCall~\cite{SymPy.jl}. \\

\href{https://mathics.github.io/}{\textbf{Mathics}} &
  A free, online CAS featuring Mathematica compatible
  syntax and functions~\cite{Mathics}. \\

\href{http://mathpix.com/}{\textbf{Mathpix}} & An iOS App that detects handwritten math as input and uses
  SymPy Gamma to evaluate the math input and generate the relevant
  steps to solve the problem~\cite{Mathpix}. \\

\href{http://openrave.org/docs/latest_stable/openravepy/ikfast/}{\textbf{IKFast}} &
  A robot kinematics compiler provided by
  \href{http://openrave.org/}{OpenRAVE}~\cite{diankov2010ikfast}. \\

\href{http://www.sagemath.org/}{\textbf{SageMath}} &
  A free open-source mathematics software system, which builds on top of many
  existing open-source packages, including SymPy~\cite{sagemath}. \\

\href{http://www.pydy.org/}{\textbf{PyDy}} & Multibody Dynamics with
  Python~\cite{gede2013constrained}. \\

\href{https://github.com/brombo/galgebra}{\textbf{galgebra}} &
  A module for geometric algebra (previously \texttt{sympy.galgebra})~\cite{galgebra}. \\

\href{http://yt-project.org/}{\textbf{yt}} & A Python package for
  analyzing and visualizing volumetric data~\cite{2011ApJS..192....9T}. \\

\href{http://sfepy.org/}{\textbf{SfePy}} &
  A Python package for solving partial
  differential equations (PDEs) in 1D, 2D, and 3D by the finite element (FE)
  method~\cite{Zienkiewicz2013FEM,cimrman2014sfepy}. \\

\href{http://quameon.sourceforge.net/}{\textbf{Quameon}} & Quantum
  Monte Carlo in Python~\cite{quameon}. \\

\href{http://lcapy.elec.canterbury.ac.nz/}{\textbf{Lcapy}} &
  An experimental Python package for teaching linear circuit analysis~\cite{lcapy}. \\
\bottomrule
\end{longtable}