### 2016-09-28 coverletter.tex

\documentclass{letter}

\usepackage{graphicx}
\usepackage{fancyhdr}
\usepackage{geometry}

\usepackage{hyperref}
% \signature{Aaron Meurer}
\address{Aaron Meurer \\
ERGS: \url{http://www.ergs.sc.edu/} \\
Nuclear Eng., Mechanical Eng. Dept. \\
University of South Carolina \\
541 Main Street, St. 009 \\
Columbia, SC 29201 \\
\url{asmeurer@gmail.com}}

\date{September 31, 2016} % stand-in

\geometry{headheight = 0.6in}
\fancypagestyle{empty}{\fancyhf{}\fancyhead[L]{\includegraphics[height=1in, keepaspectratio=true]{logo-curves.png}}}

\begin{document}

\begin{letter}{}
\opening{Dear Editors,}

We thank you and the referees for your thoughtful comments and suggestions on
our manuscript, ``SymPy: Symbolic Computing in Python''. We have endeavored to
address your comments and concerns point by point below. We give your
and the reviewers' comments on white background and our replies on gray
background.

Reviewer 1 had concerns that the paper was out of scope, but the editor has
confirmed that it is in scope. Many of the points that reviewer 3 raised were
in regards to the relevance of various sections. The intended scope of this
paper was to be an architecture paper for SymPy's 1.0 release. We have
examined reviewer 3's points through this lens. To address this, we have
restructured the manuscript to make the paper more readable to those not
already experts in SymPy or symbolic mathematics.

We thank you again for your consideration. Given these edits, the manuscript
has been greatly improved and we believe that it is now suitable for
publication in PeerJ Computer Science. If there are further questions or
concerns please do not hesitate to contact us.

% Main text
\closing{Best regards,\\
\fromsig{\includegraphics[scale=0.07]{aaron-signature.png}} \\
\fromname{Aaron Meurer}
}

\end{letter}
\end{document}