### 2016-09-29 gamma.tex
Add link to gamma page for tan(x)

\href{http://sympygamma.com}{SymPy Gamma} is a simple web application that
runs on the Google App Engine~\cite{ciurana2009google}. It executes and
displays the results of SymPy expressions as well as additional related
computations, in a fashion similar to that of Wolfram\textbar{}Alpha. For
instance, entering an integer will display its prime factors, digits in the
base-10 expansion, and a factorization diagram. Entering a function will
display its docstring; in general, entering an arbitrary expression will
display its derivative, integral, series expansion, plot, and roots.

SymPy Gamma also has several features beyond just computing the
results using SymPy.

\begin{itemize}
\item
  SymPy Gamma displays integration and differentiation steps in detail, which
  can be viewed in Figure~\ref{fig:integralsteps}.\par
  {
    \centering
    \includegraphics[width=0.7\textwidth]{integral_steps.png}
    \captionof{figure}{\href{http://www.sympygamma.com/input/?i=tan(x)}{Integral steps of $\tan (x)$}}\label{fig:integralsteps}
    \par
  }
\item
  SymPy Gamma displays the factor tree diagrams for different numbers.
\item
  SymPy Gamma saves user search queries, and offers many such similar features
  for free, which Wolfram\textbar{}Alpha only offers to its paid users.
\end{itemize}
Every input query from the user on SymPy Gamma is first parsed by its
own parser capable of handling several different forms of function names
which SymPy as a library does not support. For instance, SymPy Gamma
supports queries like \texttt{sin\ x}, whereas SymPy will only recognise
\verb|sin(x)|.

This parser converts the input query to the equivalent SymPy readable code,
which is then processed by SymPy, and the result is finally printed
with the built-in \LaTeX{} output and rendered by the SymPy Gamma web application.