### 2016-10-05 polys.tex

% Polynomials

SymPy implements a suite of algorithms for polynomial manipulation,
which ranges from relatively simple algorithms for doing arithmetic of
polynomials, to advanced methods for factoring multivariate polynomials
into irreducibles, symbolically determining real and complex root isolation
intervals, or computing Gr\"{o}bner bases.

Polynomial manipulation is useful in its own right. Within SymPy, though, it is mostly used
indirectly as a tool in other areas of the library. In fact, many mathematical
problems in symbolic computing are first expressed using entities from the
symbolic core, preprocessed, and then transformed into a problem in the
polynomial algebra, where generic and efficient algorithms are used to solve
the problem.  The solutions to the original problem are subsequently recovered from the results.
This is a common scheme in symbolic integration or summation algorithms.

SymPy implements dense and sparse polynomial representations.\footnote{In a
dense representation, the coefficients for all terms up to the degree of each
variable are stored in memory. In a sparse representation, only the nonzero
coefficients are stored.} Both are used in the univariate and multivariate
cases. The dense representation is the default for univariate polynomials. For
multivariate polynomials, the choice of representation is based on the
application. The most common case for the sparse representation is algorithms
for computing Gr\"{o}bner bases (Buchberger, F4, and
F5)~\cite{Buchberger1965thesis,Faugere1999f4,Faugere2002f5}. This is because
different monomial orderings can be expressed easily in this representation.
However, algorithms for computing multivariate GCDs or factorizations, at
least those currently implemented in SymPy~\cite{paprocki2010thesis},
are better expressed when the representation is dense. The dense multivariate
representation is specifically a recursively-dense representation, where
polynomials in $K[x_0, x_1, \dotsc, x_n]$ are viewed as a polynomials in
$K[x_0][x_1]\dotso[x_n]$. Note that despite this, the coefficient domain $K$,
can be a multivariate polynomial domain as well. The dense recursive
representation in Python gets inefficient as the number of variables increases.

Some examples for the \texttt{sympy.polys} submodule can be found in the
supplement.